---
title: "Did the Minimum Wage Kill Jobs? Replicating Card & Krueger (1994)"
subtitle: "A Difference-in-Differences Analysis of Fast-Food Employment"
description: "Replicating the key findings from Card & Krueger (1994) — a landmark study showing that New Jersey's 1992 minimum wage increase did not reduce fast-food employment."
author: "Alex Zhong"
date: "2026-04-21"
categories: [economics, difference-in-differences, R, causal inference]
format:
  html:
    toc: true
    toc-depth: 2
    toc-location: left
    code-tools: true
    fig-width: 9
    fig-height: 5.5
    fig-align: center
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
library(tidyverse)
library(gt)
library(broom)
library(scales)
library(sandwich)
library(lmtest)

theme_set(
  theme_minimal(base_size = 12) +
    theme(
      plot.title     = element_text(face = "bold", size = 13),
      plot.subtitle  = element_text(color = "grey40", size = 11),
      plot.caption   = element_text(color = "grey50", size = 9, hjust = 0),
      panel.grid.minor = element_blank(),
      strip.text     = element_text(face = "bold")
    )
)

col_nj <- "#2C5F7F"
col_pa <- "#C94F4F"
```

## Introduction

In April 1992, New Jersey raised its state minimum wage from $4.25 to $5.05 per hour — a 19% increase. The textbook prediction from a competitive labor market model is clear: a wage floor set above the equilibrium rate should reduce employment. Employers facing higher mandatory wages will hire fewer workers.

Card and Krueger (1994) challenged this prediction directly. They surveyed 410 fast-food restaurants in New Jersey and neighboring Pennsylvania in February–March 1992 (before the increase) and again in November–December 1992 (after). Pennsylvania's minimum wage remained at $4.25 throughout, making it a natural control group: geographically adjacent, economically similar, and unaffected by the policy change.

By comparing the change in employment across the two states — a technique now universally called difference-in-differences (DiD) — they identified the causal effect of the minimum wage, netting out any common trends in the fast-food industry. Their headline finding: NJ employment rose by roughly 2.76 FTE workers per store relative to PA, flatly contradicting the competitive model's prediction.

The paper ignited a debate that reshaped labor economics. This post replicates its three core empirical results: Table 2 (descriptive statistics), Table 3 (the DiD estimate), and Table 4 column (i) (the regression version), and adds a graphical look at the wage distribution that the paper covers only in tables.

## Data

The data are publicly available from David Card's website at UC Berkeley.

```{r load-data}
url     <- "https://davidcard.berkeley.edu/data_sets/njmin.zip"
tmp_zip <- tempfile(fileext = ".zip")
tmp_dir <- file.path(tempdir(), "njmin_ck")
dir.create(tmp_dir, showWarnings = FALSE, recursive = TRUE)

download.file(url, tmp_zip, mode = "wb", quiet = TRUE)
unzip(tmp_zip, exdir = tmp_dir, overwrite = TRUE)

dat_file <- list.files(tmp_dir, pattern = "\\.dat$", full.names = TRUE)[1]

var_names <- c(
  "sheet", "chain", "co_owned", "state",
  "southj", "centralj", "northj", "pa1", "pa2", "shore",
  "ncalls", "empft", "emppt", "nmgrs", "wage_st",
  "inctime", "firstinc", "bonus", "pctaff", "meal",
  "open", "hrsopen", "psoda", "pfry", "pentree",
  "nregs", "nregs11",
  "type2", "status2", "date2",
  "empft2", "emppt2", "nmgrs2", "wage_st2",
  "inctime2", "firstin2", "bonus2", "pctaff2", "meal2",
  "open2r", "hrsopen2", "psoda2", "pfry2", "pentree2",
  "nregs2", "nregs112"
)

ck_raw <- readr::read_table(
  dat_file,
  col_names = var_names,
  col_types = readr::cols(.default = readr::col_double()),
  na        = c(".", ""),
  show_col_types = FALSE
)

ck <- ck_raw |>
  as_tibble() |>
  mutate(
    state_name = if_else(state == 1, "NJ", "PA"),
    fte1 = empft + 0.5 * emppt + nmgrs,
    fte2 = case_when(
      status2 %in% c(2, 3) ~ 0,                             # closed/renovating → FTE = 0
      !is.na(empft2)        ~ empft2 + 0.5 * emppt2 + nmgrs2, # open & interviewed → compute FTE
      TRUE                  ~ NA_real_                        # refused / not reached → exclude
    ),,
    dfte = fte2 - fte1
  )

balanced <- ck |> filter(!is.na(fte1), !is.na(fte2))
```

The full dataset has `r nrow(ck)` stores (`r sum(ck$state == 1)` in NJ, `r sum(ck$state == 0)` in PA). After restricting to stores with usable employment data in both survey waves, we have `r nrow(balanced)` stores.

**Handling of closed restaurants.** Following Card & Krueger's Table 3 Column (i) specification, permanently closed stores (`status2 == 2`) and stores under renovation (`status2 == 3`) are assigned FTE = 0 in wave 2. Stores under renovation (`status2 == 3`) are also assigned FTE = 0, following the paper's convention. Stores that refused or could not be reached are treated as missing and excluded from the balanced panel.

## Replication: Table 2 — Descriptive Statistics {#table2}

Table 2 in the paper reports means and standard errors for key variables by state.

```{r table2}
smry <- balanced |>
  group_by(state_name) |>
  summarise(
    fte1_m  = mean(fte1,    na.rm = TRUE),
    fte1_se = sd(fte1,      na.rm = TRUE) / sqrt(n()),
    fte2_m  = mean(fte2,    na.rm = TRUE),
    fte2_se = sd(fte2,      na.rm = TRUE) / sqrt(n()),
    w1_m    = mean(wage_st, na.rm = TRUE),
    w1_se   = sd(wage_st,   na.rm = TRUE) / sqrt(n()),
    w2_m    = mean(wage_st2,na.rm = TRUE),
    w2_se   = sd(wage_st2,  na.rm = TRUE) / sqrt(n()),
    n       = n()
  )

nj_s <- smry |> filter(state_name == "NJ")
pa_s <- smry |> filter(state_name == "PA")

t1 <- t.test(fte1    ~ state, data = balanced)
t2 <- t.test(fte2    ~ state, data = balanced)
t3 <- t.test(wage_st ~ state, data = balanced)
t4 <- t.test(wage_st2~ state, data = balanced)

fmt <- function(m, se) sprintf("%.2f (%.2f)", m, se)

tibble(
  Variable = c(
    "FTE employment — wave 1 (before)",
    "FTE employment — wave 2 (after)",
    "Starting wage — wave 1 ($/hr)",
    "Starting wage — wave 2 ($/hr)"
  ),
  `NJ` = c(
    fmt(nj_s$fte1_m, nj_s$fte1_se),
    fmt(nj_s$fte2_m, nj_s$fte2_se),
    fmt(nj_s$w1_m,   nj_s$w1_se),
    fmt(nj_s$w2_m,   nj_s$w2_se)
  ),
  `PA` = c(
    fmt(pa_s$fte1_m, pa_s$fte1_se),
    fmt(pa_s$fte2_m, pa_s$fte2_se),
    fmt(pa_s$w1_m,   pa_s$w1_se),
    fmt(pa_s$w2_m,   pa_s$w2_se)
  ),
  `t-stat (NJ vs PA)` = c(
    sprintf("%.2f", abs(t1$statistic)),
    sprintf("%.2f", abs(t2$statistic)),
    sprintf("%.2f", abs(t3$statistic)),
    sprintf("%.2f", abs(t4$statistic))
  )
) |>
  gt() |>
  tab_header(
    title    = "Table 2: Means of Key Variables",
    subtitle = "Standard errors in parentheses. Balanced panel."
  ) |>
  tab_source_note("FTE = full-time employees + 0.5 × part-time employees + managers.") |>
  tab_source_note("Source: replication of Card & Krueger (1994), Table 2 (selected rows).") |>
  cols_label(`t-stat (NJ vs PA)` = md("*t*-stat"))
```

The pre-period NJ and PA starting wages are essentially identical (~$4.25), confirming both groups were near the binding federal minimum before the experiment. The wave-2 NJ wage jumps sharply, reflecting mandatory compliance with the new $5.05 floor.

## Replication: Table 3 — The DiD Estimate {#table3}

Table 3 is the heart of the paper. It presents average FTE employment before and after the minimum wage increase, separately for NJ and PA, and computes the difference-in-differences.

```{r table3}
m <- balanced |>
  group_by(state_name) |>
  summarise(
    before    = mean(fte1,  na.rm = TRUE),
    before_se = sd(fte1,    na.rm = TRUE) / sqrt(n()),
    after     = mean(fte2,  na.rm = TRUE)
    after_se  = sd(fte2,    na.rm = TRUE) / sqrt(n()),
    change    = mean(dfte,  na.rm = TRUE),
    change_se = sd(dfte,    na.rm = TRUE) / sqrt(n())
  )

nj_m <- m |> filter(state_name == "NJ")
pa_m <- m |> filter(state_name == "PA")

diff_before    <- nj_m$before - pa_m$before
diff_before_se <- sqrt(nj_m$before_se^2 + pa_m$before_se^2)
diff_after     <- nj_m$after  - pa_m$after
diff_after_se  <- sqrt(nj_m$after_se^2  + pa_m$after_se^2)
did            <- nj_m$change - pa_m$change
did_se         <- sqrt(nj_m$change_se^2 + pa_m$change_se^2)

tibble(
  ` ` = c("FTE employment before", "FTE employment after", "Change in FTE employment"),
  `New Jersey`   = c(fmt(nj_m$before, nj_m$before_se),
                     fmt(nj_m$after,  nj_m$after_se),
                     fmt(nj_m$change, nj_m$change_se)),
  `Pennsylvania` = c(fmt(pa_m$before, pa_m$before_se),
                     fmt(pa_m$after,  pa_m$after_se),
                     fmt(pa_m$change, pa_m$change_se)),
  `NJ − PA`      = c(fmt(diff_before, diff_before_se),
                     fmt(diff_after,  diff_after_se),
                     fmt(did,         did_se))
) |>
  gt() |>
  tab_header(
    title    = "Table 3: Average FTE Employment Before and After the Minimum Wage Increase",
    subtitle = "Standard errors in parentheses. Balanced panel."
  ) |>
  tab_style(
    style     = list(cell_text(weight = "bold"), cell_fill(color = "#f0f5fa")),
    locations = cells_body(rows = 3)
  ) |>
  tab_source_note("Source: replication of Card & Krueger (1994), Table 3, Column (i).")
```

The key cell is the bottom-right: the difference-in-differences estimate of `r sprintf("%.2f", did)` FTE workers per store (paper: 2.76). Reading down the NJ − PA column:

- Before the increase, NJ employed fewer workers than PA (`r sprintf("%.2f", diff_before)` difference).
- After the increase, the gap narrows dramatically (`r sprintf("%.2f", diff_after)`).
- The change in NJ exceeded the change in PA by `r sprintf("%.2f", did)` FTE workers — the DiD estimate (paper: 2.76).

## Replication: Table 4, Column (i) — OLS Regression {#table4}

The DiD estimate has an exact regression equivalent:

$$\Delta \text{FTE}_i = \alpha + \beta \cdot \mathbf{1}[\text{NJ}_i] + \varepsilon_i$$

The intercept $\alpha$ captures the average employment change in Pennsylvania; $\beta$ is the DiD estimate.

```{r table4}
mod <- lm(dfte ~ state, data = balanced)
ct  <- coeftest(mod, vcov = vcovHC(mod, type = "HC1"))

tidy(ct) |>
  transmute(
    Term             = c("Intercept (α — PA average change)", "NJ indicator (β — DiD estimate)"),
    Estimate         = round(estimate,  2),
    `Robust SE (HC1)`= round(std.error, 2),
    `t-stat`         = round(statistic, 2),
    `p-value`        = round(p.value,   3)
  ) |>
  gt() |>
  tab_header(
    title    = "Table 4, Column (i): OLS Regression of Change in FTE Employment",
    subtitle = "Dependent variable: ΔFTE = FTE after − FTE before"
  ) |>
  tab_style(
    style     = list(cell_text(weight = "bold"), cell_fill(color = "#f0f5fa")),
    locations = cells_body(rows = 2)
  ) |>
  tab_source_note(
    sprintf("N = %d stores.  R² = %.3f.", nrow(balanced), summary(mod)$r.squared)
  ) |>
  tab_source_note("Source: replication of Card & Krueger (1994), Table 4, Column (i).")
```

The NJ coefficient of `r round(coef(mod)[2], 2)` closely matches the paper's reported 2.76 (paper SE = 1.07).

## Additional Analysis: Did the Treatment Actually Happen? {#additional}

Causal identification with DiD requires the treatment to have actually occurred. The plot below shows the full distribution of starting wages in NJ and PA before and after the April 1992 increase — something the paper only summarizes as column means.

```{r wage-distribution}
wage_long <- balanced |>
  select(state_name, wage_st, wage_st2) |>
  pivot_longer(c(wage_st, wage_st2), names_to = "wave", values_to = "wage") |>
  mutate(
    wave = if_else(wave == "wage_st",
                   "Wave 1 — Before  (Feb–Mar 1992)",
                   "Wave 2 — After   (Nov–Dec 1992)")
  ) |>
  filter(!is.na(wage))

ggplot(wage_long, aes(x = wage, fill = state_name)) +
  geom_histogram(binwidth = 0.10, position = "identity", alpha = 0.65, color = "white") +
  geom_vline(xintercept = 4.25, linetype = "dashed", color = "grey45", linewidth = 0.7) +
  geom_vline(xintercept = 5.05, linetype = "dashed", color = col_pa,   linewidth = 0.7) +
  annotate("text", x = 4.25, y = Inf, label = "$4.25\n(old federal min)",
           vjust = 1.6, hjust = 1.1,   size = 3, color = "grey40") +
  annotate("text", x = 5.05, y = Inf, label = "$5.05\n(NJ new min)",
           vjust = 1.6, hjust = -0.08, size = 3, color = col_pa) +
  scale_fill_manual(values = c("NJ" = col_nj, "PA" = col_pa),
                    labels = c("NJ" = "New Jersey", "PA" = "Pennsylvania")) +
  scale_x_continuous(labels = dollar_format(), breaks = seq(4.0, 6.5, 0.25),
                     limits = c(3.9, 6.6)) +
  facet_wrap(~wave, ncol = 1) +
  labs(
    title    = "Starting Wage Distributions in NJ and PA, Before and After the Minimum Wage Increase",
    subtitle = "New Jersey raised its minimum wage from $4.25 to $5.05 in April 1992; Pennsylvania did not change",
    x = "Starting Wage ($/hr)", y = "Number of Stores", fill = NULL,
    caption = "Source: Card & Krueger (1994). Dashed lines mark the old ($4.25) and new NJ ($5.05) minimum wages."
  ) +
  theme(legend.position = "top")
```

- **Before:** both states cluster at $4.25 — the distributions are nearly identical.
- **After:** NJ shows a sharp spike exactly at $5.05, as employers raised wages to comply. PA is unchanged.

This confirms the natural experiment worked as intended.

## Discussion

### Replication fidelity

The replication recovers the paper's central finding: the DiD estimate is `r sprintf("%.2f", did)` FTE workers per store (paper: 2.76). Minor discrepancies can arise from two sources: (1) sample definition — the paper presents several cuts of the data with slightly different inclusion rules; (2) standard error computation — HC1 heteroskedasticity-robust standard errors are used throughout, consistent with the paper's reported SEs.

### Why DiD enables causal inference

A standard OLS regression of wave-2 employment on a New Jersey indicator would be hopelessly confounded: NJ and PA restaurants differ in chain mix, local market conditions, and countless other factors that also affect employment. DiD solves this by differencing away two layers of confounding:

1. **Time-invariant store heterogeneity** — by looking at the *change* in employment ($\Delta$FTE) rather than its level, any fixed characteristic of a store (location, chain, management quality) cancels out.
2. **Common time trends** — by subtracting Pennsylvania's average change from New Jersey's, any shock that hit the entire fast-food industry in 1992 (a national recession, changing consumer tastes) is netted out.

What remains is the *differential* change experienced by NJ restaurants relative to PA — the only difference between the two groups being New Jersey's minimum wage increase. Under the identification assumption, this differential is the causal effect of the policy.

### The parallel trends assumption

DiD's validity rests on a single key assumption: **parallel trends** — in the absence of the minimum wage increase, NJ and PA employment would have evolved along the same trajectory. This is not directly testable (we never observe the NJ counterfactual), but several pieces of evidence support it:

- **Geographic proximity:** NJ and PA share a border; the two states face similar labor demand conditions in the Philadelphia and New York metro areas.
- **Pre-period wage parity:** As shown in Table 2, wave-1 starting wages in NJ and PA are nearly identical (~$4.25), confirming both groups were at the same point in the wage distribution before the experiment.
- **Treatment validation:** The wage distribution plot shows the NJ distribution shifting sharply to $5.05 after April 1992 while PA is unchanged — the treatment occurred exactly as designed.

If parallel trends holds, the DiD estimate of `r sprintf("%.2f", did)` FTE workers per store (paper: 2.76) is a credible causal estimate, not a correlation. Card and Krueger's provocative finding — that a minimum wage increase *raised* relative employment — has since been replicated, debated, and refined over three decades, but the methodological template they established remains the gold standard for policy evaluation with natural experiments.
