---
title: "MaxDiff Analysis of MSBA Class Preferences"
subtitle: "Counts, MLE, and Bayesian estimation of the MNL model"
description: "How do MSBA students rank the ten core Rady classes? I estimate preferences from a Best–Worst Scaling survey three ways: a simple counts score, an MNL fit by maximum likelihood with a hand-coded log-likelihood and optim(), and the same MNL fit by a hand-written random-walk Metropolis–Hastings sampler."
author: "Alex Zhong"
date: "2026-05-21"
categories: [MaxDiff, MNL, MLE, Bayesian, MCMC, R, marketing analytics]
format:
  html:
    toc: true
    toc-depth: 2
    toc-location: left
    code-fold: true
    code-tools: true
    code-summary: "Show code"
    fig-width: 8
    fig-height: 5
    fig-align: center
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(purrr)
library(knitr)
library(kableExtra)
library(patchwork)

set.seed(42)              # reproducibility for the bootstrap and the MCMC

# A fixed colour for each class, so a single course can be traced across every
# chart in this document.
class_palette <- c(
  "495 Marketing Analytics" = "#1b9e77",
  "453 Business Analytics"  = "#d95f02",
  "455 Customer Analytics"  = "#7570b3",
  "452 Large Data"          = "#e7298a",
  "454 Capstone"            = "#66a61e",
  "464 SQL"                 = "#e6ab02",
  "457 BI Systems"          = "#a6761d",
  "403 AI/Math/Prog"        = "#666666",
  "451 Mktg/Fin/Ops"        = "#1f78b4",
  "444 Consulting"          = "#b2182b"
)
```


## Introduction

This report measures how strongly MSBA students at UCSD Rady prefer one core
class over another. The data come from a **MaxDiff** survey (also called
*Best–Worst Scaling*). On each screen the respondent sees a small subset of
classes — here, four — and is asked to pick the **single most-preferred** and
the **single least-preferred** class in that set. Repeating this over many
screens, and across many respondents, builds up a detailed picture of relative
preference.

Why MaxDiff rather than, say, asking people to rate each class 1–10? Humans are
reliable at *trade-offs* — "this one is better than that one" — but unreliable
at calibrating an absolute numeric scale. One respondent's "7" is another's
"9", some people never use the extremes, and ratings drift over the course of a
survey. By forcing a best and a worst pick out of a small set, MaxDiff removes
scale-use bias and makes everyone genuinely differentiate between the items.

The ten items here are the ten core MSBA classes. I estimate students' relative
preference for them three ways, and compare what each method tells us:

1. A simple **counts analysis** — the % of times a class is picked *best* minus
   the % of times it is picked *worst*.
2. The **multinomial logit (MNL)** model fit by **maximum likelihood** (MLE),
   which models the best/worst choices explicitly and gives classical standard
   errors.
3. The same MNL model fit by **Bayesian** methods, via a random-walk
   Metropolis–Hastings sampler, which gives a full posterior distribution and
   therefore easy uncertainty intervals on any derived quantity.

I expect the three methods to **agree on the broad ranking** — the best classes
will look best and the worst will look worst under all three. Where they differ
is in *how* they quantify preference and *how* they express uncertainty: counts
gives a quick descriptive score, MLE gives point estimates with standard errors,
and Bayes gives credible intervals on the things we actually care about, such as
shares of preference.


## The Data

```{r load-data}
raw <- read_csv("maxdiff_design_and_choices.csv", show_col_types = FALSE)
names(raw) <- c("rid", "task", "pos", "item", "resp")

labels_raw <- read_csv("maxdiff_item_labels.csv", show_col_types = FALSE)
item_labels <- labels_raw[, 1:2]
names(item_labels) <- c("item", "label")
item_labels <- item_labels |> filter(!is.na(item))

# Short, plot-friendly names for the ten classes, in item-number order.
short_label <- c(
  "403 AI/Math/Prog", "464 SQL", "451 Mktg/Fin/Ops", "452 Large Data",
  "453 Business Analytics", "444 Consulting", "455 Customer Analytics",
  "454 Capstone", "495 Marketing Analytics", "457 BI Systems"
)
item_labels <- item_labels |> mutate(short = short_label[item])
```

A quick word on **data cleaning**. The raw export actually contains two kinds of
screen per respondent. Most are the four-item MaxDiff screens this assignment is
about. But there is also a second block of two-row screens that pit a single
class against a phantom "Item 11" (an item that does not appear in the label
file). Those two-row screens are an *anchoring* question — a separate "would you
actually take this class?" threshold task — and they are **not** the MaxDiff
best/worst design described here. I therefore keep only the genuine four-item
MaxDiff screens, which are exactly the screens that contain one "best", one
"worst", and two unchosen items drawn from the ten real classes.

```{r filter-maxdiff}
# A genuine MaxDiff screen has exactly four rows. Keep those task; drop the
# two-row "Item 11" anchoring screens, which are a different question type.
task_sizes <- raw |> count(rid, task, name = "k")
md <- raw |>
  inner_join(filter(task_sizes, k == 4) |> select(rid, task),
             by = c("rid", "task"))
```

```{r descriptives}
N <- n_distinct(md$rid)                    # number of respondents
T <- nrow(md) / 4 / N                       # MaxDiff tasks per respondent
J <- md |> count(rid, task) |> pull(n) |> unique()   # items per task
n_items <- n_distinct(md$item)              # total classes in the study

tibble(
  Quantity = c("Respondents (N)", "MaxDiff tasks per respondent (T)",
               "Items per task (J)", "Total items in study",
               "Total MaxDiff tasks", "Total MaxDiff rows"),
  Value = c(N, T, paste(J, collapse = ", "), n_items, N * T, nrow(md))
) |>
  kable(caption = "Basic descriptives of the MaxDiff data (four-item screens only)") |>
  kable_styling(full_width = FALSE)
```

So we have **`r N` respondents**, each completing **`r T` MaxDiff screens**, with
**`r paste(J, collapse=", ")` items per screen**, drawn from **`r n_items`
classes** — matching the design exactly.

**Sanity check.** Every task should have exactly one row marked best (`+1`),
exactly one marked worst (`-1`), and the rest `0`. The code below counts any
task that violates this; the count should be zero.

```{r sanity-check}
violations <- md |>
  group_by(rid, task) |>
  summarise(n_best  = sum(resp ==  1),
            n_worst = sum(resp == -1),
            n_zero  = sum(resp ==  0),
            n_distinct_items = n_distinct(item),
            .groups = "drop") |>
  filter(!(n_best == 1 & n_worst == 1 & n_zero == 2 & n_distinct_items == 4))

cat("Tasks violating the one-best / one-worst / four-distinct rule:",
    nrow(violations), "out of", N * T, "\n")
```

**Exposure balance.** A good MaxDiff design shows each item roughly the same
number of times, so that no class has an unfair count advantage. The exposures
below are all close to `r round(nrow(md)/n_items)`, confirming a balanced
design.

```{r exposure}
md |>
  count(item, name = "times_shown") |>
  left_join(item_labels, by = "item") |>
  select(item, short, times_shown) |>
  arrange(item) |>
  kable(col.names = c("Item", "Class", "Times shown"),
        caption = "Each class is shown a roughly equal number of times") |>
  kable_styling(full_width = FALSE)
```


## Counts Analysis

The **counts analysis** is the simplest way to score the classes. For each class
$j$ we look only at the screens where it appeared, and compute the share of
those appearances on which it was picked *best* and the share on which it was
picked *worst*. The difference is a one-number score:

$$
\text{score}_j \;=\; \%\text{best}_j \;-\; \%\text{worst}_j .
$$

A score near $+1$ means a class is almost always chosen best when it is shown;
near $-1$ means it is almost always the worst; near $0$ means it is either
middle-of-the-pack or polarizing (picked best and worst about equally often).
It uses no model — just tallies — which is exactly why it is a good first look.

```{r counts-table}
counts <- md |>
  group_by(item) |>
  summarise(shown = n(),
            best  = sum(resp ==  1),
            worst = sum(resp == -1),
            .groups = "drop") |>
  mutate(pct_best  = best  / shown,
         pct_worst = worst / shown,
         score     = pct_best - pct_worst) |>
  left_join(item_labels, by = "item") |>
  arrange(desc(score)) |>
  mutate(counts_rank = row_number())

counts |>
  select(item, label, shown, best, worst, pct_best, pct_worst, score) |>
  mutate(across(c(pct_best, pct_worst, score), ~round(.x, 3))) |>
  kable(col.names = c("Item", "Class", "Shown", "Best", "Worst",
                      "% best", "% worst", "Score"),
        caption = "Counts analysis, ranked from most to least preferred") |>
  kable_styling(full_width = FALSE)
```

**Bootstrap confidence intervals.** To attach uncertainty to these scores I
resample *respondents* (not individual rows) with replacement 1,000 times,
recompute every class's score on each resample, and take the 2.5th and 97.5th
percentiles. Resampling whole respondents respects the fact that one person
contributes many correlated screens.

```{r bootstrap}
B <- 1000
md_by_resp <- split(md, md$rid)            # pre-split for speed
respondents <- names(md_by_resp)

boot_scores <- matrix(NA_real_, nrow = B, ncol = n_items)
for (b in seq_len(B)) {
  samp <- sample(respondents, length(respondents), replace = TRUE)
  d    <- bind_rows(md_by_resp[samp])
  s <- d |>
    group_by(item) |>
    summarise(score = sum(resp == 1)/n() - sum(resp == -1)/n(), .groups = "drop")
  boot_scores[b, s$item] <- s$score
}

boot_ci <- apply(boot_scores, 2, quantile, probs = c(.025, .975), na.rm = TRUE)
counts <- counts |>
  mutate(ci_lo = boot_ci[1, item],
         ci_hi = boot_ci[2, item])
```

```{r counts-plot}
#| fig-cap: "Counts score (% best − % worst) for each class, with 95% bootstrap CIs."
ggplot(counts, aes(x = score, y = reorder(short, score), fill = short)) +
  geom_col(width = 0.7) +
  geom_errorbarh(aes(xmin = ci_lo, xmax = ci_hi), height = 0.3, colour = "grey25") +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
  scale_fill_manual(values = class_palette, guide = "none") +
  labs(x = "Counts score  (% best − % worst)", y = NULL,
       title = "Counts analysis of MSBA class preferences") +
  theme_minimal(base_size = 12)
```

**What we see.** *MGTA 495 Marketing Analytics* is the runaway favourite — its
score (≈ `r round(counts$score[counts$item==9],2)`) sits far above every other
class and its confidence interval does not come close to overlapping the rest.
A cluster of analytics-heavy electives — *453 Business Analytics*, *455 Customer
Analytics*, and *452 Large Data* — form a clear second tier with positive
scores. The least-preferred classes are *444 Analytics Consulting*, *451
Marketing/Finance/Operations*, and *403 AI Math/Prog*, all with significantly
negative scores. The middle classes (*454 Capstone*, *464 SQL*) have scores
near zero with intervals straddling it, meaning the data cannot confidently
place them above or below average. The one mild surprise is that the *403 AI
Math/Programming* foundation course lands near the bottom — perhaps because it
is a required summer prerequisite rather than a chosen elective.


## From MaxDiff Data to MNL Choices

The counts analysis throws away information: it treats each screen's best pick
and worst pick in isolation and never uses the fact that an item was *available
but not chosen*. The **multinomial logit (MNL)** model fixes this by modelling
the choice process itself. (This is the conjoint/MNL framework from the week-5
deck.)

We give each class $j$ a latent **utility coefficient** $\beta_j$. On a screen
showing the set $\{a,b,c,d\}$, the probability that item $b$ is the one picked
*best* is the soft-max of the utilities over the shown set:

$$
P(\text{best} = b) \;=\; \frac{\exp(\beta_b)}{\exp(\beta_a)+\exp(\beta_b)+\exp(\beta_c)+\exp(\beta_d)} .
$$

The respondent then picks a *worst* item from the **three that remain** after
the best is removed. Picking the worst is just picking the best of the
*negated* utilities, so

$$
P(\text{worst} = c \mid \text{best} = b) \;=\; \frac{\exp(-\beta_c)}{\exp(-\beta_a)+\exp(-\beta_c)+\exp(-\beta_d)} .
$$

So **each MaxDiff screen yields two MNL choice observations**: one best-pick from
the full set of four, and one worst-pick from the remaining three. Crucially,
every shown-but-unchosen class now contributes to the likelihood through the
denominators — that is the extra information the counts analysis ignored.

**Identifiability.** The soft-max is unchanged if we add the same constant to
every $\beta$, so only $J-1 = 9$ of the ten coefficients are identified. I fix
the first class as the reference, $\beta_1 = 0$, and estimate
$\beta_2,\dots,\beta_{10}$ relative to it. Every reported $\beta$ is therefore a
utility *relative to MGTA 403*.

```{r reshape-tasks}
# One row per task is the natural unit for the likelihood: store the four item
# numbers shown, plus which one was best and which was worst.
tasks <- md |>
  group_by(rid, task) |>
  summarise(items = list(item),
            best  = item[resp ==  1],
            worst = item[resp == -1],
            .groups = "drop")

TASK_ITEMS <- lapply(tasks$items, as.integer)   # list of length-4 integer vectors
BEST  <- as.integer(tasks$best)                 # best item per task
WORST <- as.integer(tasks$worst)                # worst item per task
n_tasks <- length(BEST)
```


## MNL via Maximum Likelihood

Summing the log-probabilities of the best and worst picks over all
$N \cdot T = `r n_tasks`$ tasks gives the log-likelihood

$$
\ell(\boldsymbol\beta) \;=\; \sum_{\text{tasks}}
\Big[\, \beta_{j^*_b} - \log\!\!\sum_{j\in\text{shown}}\!\! e^{\beta_j}
\;-\; \beta_{j^*_w} - \log\!\!\!\sum_{j\in\text{shown}\setminus\{j^*_b\}}\!\!\! e^{-\beta_j}\,\Big],
$$

where $j^*_b$ and $j^*_w$ are the best and worst items chosen in that task. The
first two terms are the best-pick log-probability; the last two are the
worst-pick log-probability, with utilities flipped in sign and the chosen-best
item removed from the set.

```{r loglik}
# Log-likelihood. `beta9` holds beta_2..beta_10; beta_1 is pinned to 0 inside.
# Written as a plain loop over tasks so the math above maps line-for-line onto
# the code. Uses log(sum(exp(.))) directly; the utilities here are small enough
# that no extra log-sum-exp shifting is needed for stability.
loglik <- function(beta9) {
  beta <- c(0, beta9)                 # length-10 utility vector, beta_1 = 0
  total <- 0
  for (t in seq_len(n_tasks)) {
    shown <- TASK_ITEMS[[t]]
    b <- BEST[t]; w <- WORST[t]
    # best pick: soft-max over all four shown utilities
    total <- total + beta[b] - log(sum(exp(beta[shown])))
    # worst pick: soft-max over the three remaining, utilities negated
    remaining <- shown[shown != b]
    total <- total - beta[w] - log(sum(exp(-beta[remaining])))
  }
  total
}
```

Maximizing with `optim` (BFGS, `fnscale = -1` to turn the minimizer into a
maximizer) and asking for the Hessian:

```{r mle-fit}
fit <- optim(par = rep(0, 9), fn = loglik,
             method = "BFGS", hessian = TRUE,
             control = list(fnscale = -1))

cat("Converged (0 = yes):", fit$convergence,
    " | log-likelihood at optimum:", round(fit$value, 2), "\n")

# Standard errors from the inverse of the negative Hessian (the observed
# information matrix). With fnscale = -1, optim returns the Hessian of the
# log-likelihood itself, so -Hessian is the (positive-definite) information.
se9 <- sqrt(diag(solve(-fit$hessian)))

beta_hat <- c(0, fit$par)
se_hat   <- c(0, se9)
```

To make the coefficients interpretable I convert them to **shares of
preference** — what fraction of choices each class would win if all ten appeared
on one screen:

$$
\hat s_j \;=\; \frac{\exp(\hat\beta_j)}{\sum_{k=1}^{10}\exp(\hat\beta_k)} .
$$

```{r mle-table}
mle <- tibble(item = 1:10, beta = beta_hat, se = se_hat) |>
  mutate(share = exp(beta) / sum(exp(beta))) |>
  left_join(item_labels, by = "item") |>
  arrange(desc(share)) |>
  mutate(mle_rank = row_number())

mle |>
  select(item, label, beta, se, share) |>
  mutate(across(c(beta, se, share), ~round(.x, 3))) |>
  kable(col.names = c("Item", "Class", "β̂ (vs MGTA 403)", "SE(β̂)", "Share"),
        caption = "MLE estimates: utility coefficients, standard errors, and shares of preference") |>
  kable_styling(full_width = FALSE)
```

```{r mle-plot}
#| fig-cap: "MLE shares of preference — the share of choices each class would win if all ten were on one screen."
ggplot(mle, aes(x = share, y = reorder(short, share), fill = short)) +
  geom_col(width = 0.7) +
  scale_fill_manual(values = class_palette, guide = "none") +
  scale_x_continuous(labels = scales::percent) +
  labs(x = "Share of preference", y = NULL,
       title = "MNL (MLE) shares of preference") +
  theme_minimal(base_size = 12)
```

**MLE vs counts.** The two rankings are nearly identical: MGTA 495 dominates,
the same analytics electives form the upper tier, and 444 Consulting / 451 MFO /
403 AI sit at the bottom under both methods. The MNL refines the middle of the
table. Because it credits every *shown-but-not-chosen* appearance as evidence of
lower utility, the MNL spreads the middle classes apart a little differently
than counts does — small reorderings among near-tied classes (e.g., 464 SQL,
454 Capstone, 457 BI) are exactly where a model that uses more of the data and a
raw tally can disagree.

```{r mle-vs-counts}
left_join(
  counts |> select(item, short, counts_rank),
  mle    |> select(item, mle_rank),
  by = "item"
) |>
  arrange(mle_rank) |>
  kable(col.names = c("Item", "Class", "Counts rank", "MLE rank"),
        caption = "Counts rank vs MLE rank") |>
  kable_styling(full_width = FALSE)
```


## MNL via Bayesian Estimation

Now I fit the *same* MNL model the Bayesian way (week-7 deck). Instead of a
single point estimate, we characterise the whole **posterior** distribution of
$\boldsymbol\beta$,

$$
\pi(\boldsymbol\beta \mid \boldsymbol y) \;\propto\; L(\boldsymbol\beta)\cdot\pi(\boldsymbol\beta),
$$

with a weakly informative prior $\boldsymbol\beta \sim N(\boldsymbol 0,\,10\cdot I)$
(an SD of $\sqrt{10}\approx 3.2$ on each coefficient — diffuse, but not absurd).
Working on the log scale, the log-posterior is the log-likelihood plus the log
of the normal prior density.

```{r log-post}
log_post <- function(beta9) {
  loglik(beta9) + sum(dnorm(beta9, mean = 0, sd = sqrt(10), log = TRUE))
}
```

**The sampler.** A textbook random-walk Metropolis: propose a small Gaussian
step from the current $\boldsymbol\beta$, and accept it with probability
$\min\!\big(1,\, e^{\,\text{logpost}^\star - \text{logpost}}\big)$ (compared on
the log scale against $\log U$). Reject and we stay put. Over many iterations the
chain spends time in each region of $\boldsymbol\beta$-space in proportion to its
posterior probability.

```{r mh-sampler}
metropolis <- function(beta0, n_iter, step) {
  K      <- length(beta0)
  draws  <- matrix(NA_real_, n_iter, K)
  beta   <- beta0
  lp     <- log_post(beta)
  accept <- 0
  for (t in seq_len(n_iter)) {
    beta_star <- beta + rnorm(K, 0, step)      # symmetric random-walk proposal
    lp_star   <- log_post(beta_star)
    if (log(runif(1)) < lp_star - lp) {        # accept/reject on the log scale
      beta <- beta_star; lp <- lp_star; accept <- accept + 1
    }
    draws[t, ] <- beta
  }
  list(draws = draws, accept_rate = accept / n_iter)
}
```

**Tuning the step size.** We want an acceptance rate of roughly 25–40%. Too big
a step and almost every proposal is rejected (the chain gets stuck); too small
and it crawls. The rule of thumb: if acceptance is too low, shrink the step; too
high, grow it. A few short trial runs locate the sweet spot:

```{r mh-tune}
tune <- map_dfr(c(0.20, 0.10, 0.07, 0.05), function(s) {
  r <- metropolis(fit$par, n_iter = 2000, step = s)
  tibble(step = s, accept_rate = round(r$accept_rate, 3))
})
kable(tune, col.names = c("Step size", "Acceptance rate"),
      caption = "Step-size tuning: a step of 0.07 lands in the 25–40% target") |>
  kable_styling(full_width = FALSE)
```

A **step of 0.07** sits squarely in the target band, so I run the real chain
with it: 20,000 iterations, discarding the first 25% (5,000) as burn-in.

```{r mh-run}
n_iter <- 20000
run <- metropolis(beta0 = rep(0, 9), n_iter = n_iter, step = 0.07)
cat("Final acceptance rate:", round(run$accept_rate, 3), "\n")

burn   <- seq_len(n_iter * 0.25)               # first 25% = burn-in
post9  <- run$draws[-burn, ]                    # post-burn-in draws, beta_2..10
post   <- cbind(0, post9)                       # prepend the reference beta_1 = 0
colnames(post) <- paste0("b", 1:10)
```

**Diagnostics.** The trace plots below should look like fuzzy caterpillars —
rapidly mixing around a stable level, with no slow drift — which confirms the
chain has converged. I show three coefficients spanning the range (the favourite
495, a mid-tier 453, and the least-liked 444).

```{r mh-trace}
#| fig-cap: "Post-burn-in trace plots for three β coefficients — fuzzy caterpillars indicate good mixing."
trace_df <- as_tibble(post[, c(9, 5, 6)]) |>
  setNames(c("β9 — 495 Marketing", "β5 — 453 Business", "β6 — 444 Consulting")) |>
  mutate(iter = row_number()) |>
  pivot_longer(-iter, names_to = "param", values_to = "value")

ggplot(trace_df, aes(iter, value)) +
  geom_line(linewidth = 0.2, colour = "grey30") +
  facet_wrap(~param, ncol = 1, scales = "free_y") +
  labs(x = "Iteration (post burn-in)", y = "β draw") +
  theme_minimal(base_size = 11)
```

**Posterior summaries.** Posterior mean, 95% credible interval, and posterior
mean share for each class. The shares are computed the easy Bayesian way: push
*every* draw through the soft-max and average the resulting shares.

```{r mh-summary}
post_mean <- colMeans(post)
post_sd   <- apply(post, 2, sd)
post_ci   <- apply(post, 2, quantile, probs = c(.025, .975))

# Shares: soft-max each draw, then average across draws.
share_draws <- t(apply(post, 1, function(b) exp(b) / sum(exp(b))))
share_mean  <- colMeans(share_draws)
share_ci    <- apply(share_draws, 2, quantile, probs = c(.025, .975))

bayes <- tibble(
  item       = 1:10,
  post_mean  = post_mean,
  ci_lo      = post_ci[1, ],
  ci_hi      = post_ci[2, ],
  post_sd    = post_sd,
  share      = share_mean,
  share_lo   = share_ci[1, ],
  share_hi   = share_ci[2, ]
) |>
  left_join(item_labels, by = "item") |>
  arrange(desc(share)) |>
  mutate(bayes_rank = row_number())

bayes |>
  mutate(CI = sprintf("[%.2f, %.2f]", ci_lo, ci_hi)) |>
  select(item, label, post_mean, CI, share) |>
  mutate(across(c(post_mean, share), ~round(.x, 3))) |>
  kable(col.names = c("Item", "Class", "Posterior mean β", "95% CI", "Posterior mean share"),
        caption = "Bayesian posterior summaries") |>
  kable_styling(full_width = FALSE)
```

```{r bayes-plot}
#| fig-cap: "Posterior mean shares of preference with 95% credible intervals."
ggplot(bayes, aes(x = share, y = reorder(short, share), fill = short)) +
  geom_col(width = 0.7) +
  geom_errorbarh(aes(xmin = share_lo, xmax = share_hi), height = 0.3, colour = "grey25") +
  scale_fill_manual(values = class_palette, guide = "none") +
  scale_x_continuous(labels = scales::percent) +
  labs(x = "Posterior mean share of preference", y = NULL,
       title = "Bayesian MNL shares of preference (95% credible intervals)") +
  theme_minimal(base_size = 12)
```

**Bayes vs MLE.** The two methods land in almost exactly the same place: the
posterior means track the MLE point estimates to two decimals, and the posterior
SDs match the MLE standard errors closely.

```{r bayes-vs-mle}
left_join(
  tibble(item = 1:10, post_mean = post_mean, post_sd = post_sd),
  tibble(item = 1:10, beta = beta_hat, se = se_hat),
  by = "item"
) |>
  left_join(item_labels |> select(item, short), by = "item") |>
  arrange(item) |>
  mutate(across(c(post_mean, post_sd, beta, se), ~round(.x, 3))) |>
  select(item, short, post_mean, beta, post_sd, se) |>
  kable(col.names = c("Item", "Class", "Bayes mean β", "MLE β̂", "Bayes SD", "MLE SE"),
        caption = "Posterior mean vs MLE estimate, posterior SD vs MLE SE") |>
  kable_styling(full_width = FALSE)
```

This agreement is exactly what theory predicts: with 680 tasks the likelihood
dominates the diffuse $N(0, 10)$ prior, so the posterior is essentially the
(normalized) likelihood and the Bayesian answer reproduces the MLE. The only
systematic difference is that the posterior SDs are a hair *smaller* than the
MLE SEs — the prior, even when weak, gently pulls extreme coefficients toward
zero and slightly reduces their variance. The real payoff of the Bayesian fit is
not the $\beta$'s themselves but the credible intervals on the **shares**: we got
those for free by pushing each draw through the soft-max, with no delta-method
approximation.


## Comparing the Three Methods

```{r compare-table}
comparison <- counts |>
  select(item, short, counts_score = score, counts_rank) |>
  left_join(mle   |> select(item, mle_share = share, mle_rank), by = "item") |>
  left_join(bayes |> select(item, bayes_share = share, bayes_rank), by = "item") |>
  arrange(mle_rank)

comparison |>
  mutate(across(c(counts_score, mle_share, bayes_share), ~round(.x, 3))) |>
  select(short, counts_score, counts_rank, mle_share, mle_rank, bayes_share, bayes_rank) |>
  kable(col.names = c("Class", "Counts score", "Counts rank",
                      "MLE share", "MLE rank", "Bayes share", "Bayes rank"),
        caption = "Side-by-side comparison of all three methods, sorted by MLE rank") |>
  kable_styling(full_width = FALSE)
```

```{r compare-plot}
#| fig-width: 12
#| fig-height: 6
#| fig-cap: "The three methods side by side. Colours are consistent, so any one class can be traced across all panels."
p_counts <- ggplot(counts, aes(reorder(short, score), score, fill = short)) +
  geom_col() + coord_flip() +
  scale_fill_manual(values = class_palette, guide = "none") +
  labs(title = "Counts", x = NULL, y = "% best − % worst") +
  theme_minimal(base_size = 11)

p_mle <- ggplot(mle, aes(reorder(short, share), share, fill = short)) +
  geom_col() + coord_flip() +
  scale_fill_manual(values = class_palette, guide = "none") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "MLE", x = NULL, y = "Share") +
  theme_minimal(base_size = 11)

p_bayes <- ggplot(bayes, aes(reorder(short, share), share, fill = short)) +
  geom_col() + coord_flip() +
  scale_fill_manual(values = class_palette, guide = "none") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Bayes", x = NULL, y = "Share") +
  theme_minimal(base_size = 11)

p_counts + p_mle + p_bayes
```

**Where they agree.** At the top and bottom the three methods are unanimous:
*MGTA 495 Marketing Analytics* is everyone's clear favourite, and *444
Consulting*, *451 MFO*, and *403 AI Math* are reliably at the bottom. This is the
usual pattern — the extremes are easy and every reasonable method recovers them.

**Where they disagree.** Any disagreement is confined to the middle of the
table, among classes whose utilities are close enough that a little sampling
noise can flip their order (e.g., the near-tied 464 SQL / 454 Capstone / 457 BI
group). These are precisely the cases where the choice of method matters least
substantively but most visibly in a ranking.

**What the MNL adds over counts.** The counts score only uses the best and worst
*picks*. The MNL additionally uses every *shown-but-not-picked* item: each
appearance in a denominator is evidence that a class lost to the others on that
screen. The model therefore extracts more information from the same data and
produces estimates with a coherent probabilistic interpretation (utilities and
shares) rather than a bare descriptive score.

**What Bayes adds over MLE.** For the $\beta$'s, almost nothing here — the
estimates coincide. The advantage is **uncertainty on derived quantities**. A
credible interval on a share of preference is trivial under Bayes: compute the
share for each MCMC draw and read off the quantiles. Getting the same interval
from the MLE requires the delta method (propagating the Hessian through the
nonlinear soft-max). The Bayesian machinery also extends naturally to richer
questions — market-simulation scenarios, "what if we added a new class?", or a
hierarchical model for *individual-level* preferences.


## Discussion

**The substantive finding.** MSBA students clearly prefer *MGTA 495 Marketing
Analytics* above all other core classes — by a wide and statistically clear
margin under every method. Behind it sits a tier of hands-on analytics electives
(*453 Business Analytics*, *455 Customer Analytics*, *452 Large Data*), while the
foundational and process-oriented courses (*444 Analytics Consulting*, *451
Marketing/Finance/Operations*, *403 AI Math/Programming*) are the least
preferred. The most plausible reading is that students gravitate toward applied,
tool-and-technique electives and rate required-foundation or
breadth/process courses lower. The mild surprise is how low the *403 AI
Math/Programming* course lands given the field's enthusiasm for AI — likely
because it is a required summer prerequisite experienced as work to get through,
not an elective chosen out of interest.

**The methodological finding** — which tool for which job:

- **Counts** — a fast, model-free descriptive summary. Reach for it when you
  need a quick read or an executive dashboard, and you are happy with a relative
  score rather than a probability.
- **MLE MNL** — rigorous point estimates with classical standard errors, and a
  proper choice model underneath. Reach for it when you need a single defensible
  number to report and trust large-sample theory for its uncertainty.
- **Bayesian MNL** — when you want uncertainty on *derived* quantities (shares,
  market-share simulations, "what if we added a class?" scenarios) without
  delta-method gymnastics, or when you want to extend the model hierarchically to
  estimate each student's individual preferences.

**Caveats.** The survey was taken by a self-selected sample of current MSBA
students, so these preferences may not generalize to other programs, future
cohorts, or prospective students. Responses can also be shaped by where a class
falls in the program sequence, which instructor teaches it, prior coursework,
and even item-presentation order within the survey. The MNL here is a single,
*aggregate* (homogeneous) model — it assumes everyone shares the same utilities,
which surely understates real heterogeneity in taste.

**What I would do next.** The natural extension is a **hierarchical (mixed) MNL**
that gives each respondent their own $\boldsymbol\beta$ drawn from a population
distribution. That would reveal whether the apparent consensus hides distinct
student segments (say, "marketing-leaning" vs "engineering-leaning" students) and
would let us simulate how preference shares shift if a new class were added or an
existing one dropped — the kind of forward-looking question the aggregate model
cannot answer.
